---
title: "Human CTB BPS Project"
author: "Sean Nguyen"
output:
  html_document: 
    df_print: paged
    theme: flatly
  html_notebook:
    df_print: paged
  pdf_document: default
---
#Overall Objective 


##Load Libraries

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom)
library(plotly)

```

##Import data
```{r, message=FALSE, warning=FALSE, include=FALSE}
#setwd("~/GitHub/Experiments/CTB_BPS/data/")
setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Experiments/CTB_BPS/data/")


raw_data <- "CTB_pilot_48hr_96hr-ExperimentSummary.csv"
key_data <- "Key_hCTBPetroff_Lopez.csv"
standard_data <- "std_CTB_pilot_48hr_96hr-ExperimentSummary.csv"


data <- read_csv(raw_data)
standards <- read_csv(standard_data)
key <- read_csv(key_data)

data
standards
key
```

##Convert data from 'wide' to 'long' format
```{r, message=FALSE, warning=FALSE}
# data
data1 <- data %>%
  gather(Sample,Count,2:49)

# Separate samples by identifiers 
data2 <- data1 %>% 
  separate(Sample, into=c("Sample_ID","Dilution_factor",
                          "Injection","Tech_rep", sep = "_")) %>% 
  select(-`_`)

# Standards
standards1 <- standards %>% 
  gather(Sample,Count,2:25)


standards2 <- standards1 %>% 
  separate(Sample, into=c("Sample_ID","When","Dilution_factor",
                          "Nano_day","Injection","Tech_Rep", sep = "_")) %>% 
  select(-`_`)


```


##Factor the data into categorical variables
```{r, warning=FALSE}
# Refactoring Columns for samples
data2$Sample_ID <- as.factor(data2$Sample_ID)
data2$Dilution_factor <- as.numeric(data2$Dilution_factor)
data2$Injection<- as.factor(data2$Injection)
data2$Tech_rep <- as.numeric(data2$Tech_rep)

data2

# Refactoring COlumns for key
key$Sample_ID <- as.factor(key$Sample_ID)
key$Time <- as.factor(key$Time)
key$Treatment <- as.factor(key$Treatment)
key$Volume <- as.numeric(key$Volume)

key$Treatment <- factor(key$Treatment,levels = c('DMSO','EGF','BPS','BPS_EGF'))

key

# Refactoring columns for standards
standards2$Sample_ID <- as.factor(standards2$Sample_ID)
standards2$When <- as.factor(standards2$When)
standards2$Dilution_factor <- as.numeric(standards2$Dilution_factor)
standards2$Injection <- as.factor(standards2$Injection)
standards2$Nano_day <- as.numeric(standards2$Nano_day)

standards2

```



###Back calculate standards
```{r}
standards2 <- standards2 %>% 
  mutate(True_Count=Dilution_factor*Count)

# Set the correct order of 'categorical factors'
standards2$Nano_day <-  factor(standards2$Nano_day, levels=c('1','2'))
standards2$When <- factor(standards2$When, levels=c('before','after'))

standards2
```

###Summarize three technical standard replicates
```{r}
standards3 <- standards2 %>% 
  group_by(particle_size,Sample_ID,When,Dilution_factor,Nano_day,Injection) %>% 
  summarise( tech_N = length(True_Count),
             tech_mean = mean(True_Count),
             tech_sd = sd(True_Count),
             tech_se = tech_sd/sqrt(tech_N))
standards3
```


###Summarize standards by injection
```{r}
standards4 <- standards3 %>% 
  group_by(Nano_day,When,particle_size) %>% 
  summarise( inj_N = length(tech_mean),
             inj_mean = mean(tech_mean),
             inj_sd = sd(tech_mean),
             inj_se = inj_sd/sqrt(inj_N))
standards4

```

###Plot before and after plots, facet by experimental day
```{r}
std_plot <- standards4 %>% 
  ggplot(aes(x = particle_size, y = inj_mean, color=When))+
  geom_line(size=2) + xlim(0,300)+ #line size, x-axis scale
  geom_ribbon(aes(ymin=inj_mean-inj_se, ymax=inj_mean+inj_se),
              alpha=0.4,fill = alpha('grey12', 0.2)) + #error bars
  scale_y_continuous(expand=c(0,0))+ #set bottom of graph
  xlab("Particle Size") + # X axis label
  ylab("\nMean Particle Concentration/ml\n") + # Y axis label
  ggtitle("Nanosight Histogram of\n100nm Standards")+ #title
  labs(color="Condition")+ #Label table title
  facet_wrap(~ Nano_day)

std_plot


# ggsave("Standards_histogram_plot.png",
#        height = 5, width = 7, dpi = 300, units= "in")
```

### Standards particle concentrations from each experimental day
```{r}
standards_df <- standards4 %>% 
  group_by(Nano_day,When) %>% 
  summarise(total=sum(inj_mean))

standards_df
```

###Bar graph of standards particle concentrations
```{r}
standards_bar <- standards_df %>% 
  ggplot(aes(x=Nano_day,y=total,fill=When))+
  geom_col(position="dodge")+
  scale_y_continuous(expand=c(0,0))+ #set bottom of graph
  xlab("Experimental Day") + # X axis label
  ylab("\nMean Particle Concentration/ml\n") + # Y axis label
  ggtitle("Nanosight Histogram of\n100nm Standards")+ #title
  labs(color="When") #Label table title

standards_bar

# ggsave("Standards_bar_plot.png",
#        height = 5, width = 7, dpi = 300, units= "in")
```

###Intraassay variability
```{r}
Intra.assay_cv <- standards_df %>% 
  group_by(Nano_day) %>% 
  summarise(Intra_Day_N = length(total),
            Intra_Day_mean = mean(total),
            Intra_Day_sd = sd(total),
            Intra_Day_se = Intra_Day_sd/sqrt(Intra_Day_N),
            Intra_Day_cv = Intra_Day_sd/Intra_Day_mean )
Intra.assay_cv

# # Save as .csv
# write_csv(Intra.assay_cv,"Intra.assay_cv.csv")
```

###Interassay variability
```{r}
Inter.assay_cv <- Intra.assay_cv %>% 
  summarise(Inter_Day_N = length(Intra_Day_mean),
            Inter_Day_mean = mean(Intra_Day_mean),
            Inter_Day_sd = sd(Intra_Day_mean),
            Inter_Day_se = Inter_Day_sd/sqrt(Inter_Day_N),
            Inter_Day_cv = Inter_Day_sd/Inter_Day_mean )
Inter.assay_cv

# # Save as .csv
# write_csv(Inter.assay_cv,"Inter.assay_cv.csv")
```



#Sample analysis

###Back calculate the original concentration of the sample
```{r}
data2 <- data2 %>% 
  mutate(True_Count = Dilution_factor*Count)
data2
```


### Average three technical readings
```{r}
data3 <- data2 %>% 
  group_by(particle_size,Sample_ID,Dilution_factor,Injection) %>% 
  summarise( tech_N = length(True_Count),
             tech_mean = mean(True_Count),
             tech_sd = sd(True_Count),
             tech_se = tech_sd/sqrt(tech_N))
data3
```

##Summarize samples by injection (average both injections)
```{r, warning=FALSE}

data4 <- data3 %>% 
  group_by(particle_size,Sample_ID,Dilution_factor) %>% 
  summarise( inj_N = length(tech_mean),
             inj_mean = mean(tech_mean),
             inj_sd = sd(tech_mean),
             inj_se = inj_sd/sqrt(inj_N))
data4

# Average technical replicates and merge with key
merge <- left_join(key,data3, by= "Sample_ID")

merge

# Save as .csv
# write_csv(merge,"Technical_replicate_average.csv")
 
# Average injection replicates and merge with key
merge1 <- left_join(key,data4, by= "Sample_ID")

merge1

# #Save as .csv
# write_csv(merge1,"Injection_replicate_average.csv")

```



#Quick visualizations

### Graphing all samples
```{r, warning=FALSE}
sample_plot <- merge %>%
  ggplot(aes(x=particle_size, y=tech_mean,color=Injection ))+ #plot
  geom_ribbon(aes(ymin=tech_mean-tech_se,
                  ymax=tech_mean+tech_se),
                  alpha=0.2,fill = alpha('grey12', 0.2)) + #error bars
  geom_line(size=2.0, alpha = 0.8) + xlim(0,500)+ #line size, x-axis scale
  scale_y_continuous(expand=c(0,0))+ #set bottom of graph
  xlab("Particle Size") + # X axis label
  ylab("\nMean Particle Concentration/ml\n") + # Y axis label
  ggtitle("Nanosight Histogram of\nhCTBS treated with BPS")+ #title
  labs(color="Injection")+ #Label table title
  facet_grid(Time ~ Treatment)
  # geom_vline(xintercept = 200)+
  # annotate("text", x= 350, y = 1E8, label= "200nm")

sample_plot

# ggsave("Nanosight_Sample_Histogram.png", plot = sample_plot,
#        height = 10, width = 14, dpi = 200, units= "in")
```
###Interactive Plot
```{r}
ggplotly(sample_plot)
```



### Particle concentration values for each of the samples
```{r}
merge2 <- merge1 %>% 
  group_by(Time, Treatment, Volume) %>% 
  summarise(particle_conc=sum(inj_mean))
merge2
```

### Correct for resuspension volume
```{r}
merge3 <- merge2 %>% 
  mutate(particle_count = (Volume/1000)*particle_conc, # Create new column with number of particles
         corrected_particle_conc = (particle_conc/1E9)) # Create new column with correct particle concentration
merge3

# Save as .csv
# write_csv(merge3,"Adjusted_particle_concentration.csv")

```


###Barplot
```{r, warning=TRUE}
plot1 <- merge3 %>% 
  ggplot(aes(x = Treatment, y = corrected_particle_conc, fill = Treatment)) +
  geom_bar(aes(text = paste("Particle Concentration:",
                            corrected_particle_conc)),
           stat="identity", position = "dodge")+
  xlab("\nTreatment\n") + # X axis label
  ylab("\nMean Vessicle Concentration\n(10^9 particles/ ml)\n") + # Y axis label
  ggtitle("Effect of BPS on Extracellular Vessicle\nRelease of hCTBs\n")+
  scale_y_continuous(breaks = seq(0,14,2),
                     limits = c(0,14),
                     expand = c(0,0))+ # set bottom of graph
  labs(color="Condition")+ # Label table title
  facet_wrap(~Time)

  
plot1

# ggsave("BPS_treated_hCTBs_48_96_facet_plot.png",
#       height = 8, width = 10, dpi = 600, units= "in")
```
##Interactive Plot
```{r}
ggplotly(plot1)
```

###Barplot
```{r}
plot2 <- merge3 %>%
  group_by(Time) %>% 
  ggplot(aes(x = Treatment, y = corrected_particle_conc, fill = Time )) +
  geom_bar(position = "dodge", stat = "identity")+
  xlab("\nTreatment\n") + # X axis label
  ylab("\nMean Vessicle Concentration\n(10^9 particles/ ml)\n") + # Y axis label
  ggtitle("Effect of BPS on Extracellular Vessicle\nRelease of hCTBs\n")+
  scale_y_continuous(breaks = seq(0,14,2),
                     limits = c(0,14),
                     expand = c(0,0))+ # set bottom of graph
  labs(fill= "Time (hr)")

plot2

 # ggsave("BPS_treated_hCTBs_48_96_plot.png",
 #       height = 5, width = 7, dpi = 600, units= "in")
```
## Interactive Plot
```{r}
ggplotly(plot2)
```



# Statistics

```{r}
fit <- aov(corrected_particle_conc ~ Time * Treatment ,data=merge3)

tidy(fit)
```

