---
title: "Explant Culture Analysis"
author: "Sean Nguyen"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
#Overall Objective 


##Load Libraries

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(broom)
library(plotly)
```

##Import data
```{r, message=FALSE, warning=FALSE, include=FALSE}
setwd("~/GitHub/Experiments/explant_conditions")
#setwd("~/Library/Mobile\ Documents/com~apple~CloudDocs/time-course/data")

raw_data <- "explant_combined-ExperimentSummary.csv"
key_data <- "explant_key.csv"
standard_data <- "standards_combined-ExperimentSummary.csv"


data <- read_csv(raw_data)
standards <- read_csv(standard_data)
key <- read_csv(key_data)

data
standards
key
```

##Convert data from 'wide' to 'long' format
```{r, message=FALSE, warning=FALSE}
# data
data1 <- data %>%
  gather(Sample,Count,2:37)

# Separate samples by identifiers 
data2 <- data1 %>% 
  separate(Sample, into=c("Sample_ID","Dilution_factor",
                          "Injection","Tech_rep", sep = "_")) %>% 
  select(-`_`)

# Standards
standards1 <- standards %>% 
  gather(Sample,Count,2:13)


standards2 <- standards1 %>% 
  separate(Sample, into=c("Sample_ID","When","Dilution_factor",
                          "Nano_day","Injection","Tech_Rep", sep = "_")) %>% 
  select(-`_`)


```


##Factor the data into categorical variables
```{r, warning=FALSE}
# Refactoring Columns for samples
data2$Sample_ID <- as.factor(data2$Sample_ID)
data2$Dilution_factor <- as.numeric(data2$Dilution_factor)
data2$Injection<- as.factor(data2$Injection)
data2$Tech_rep <- as.numeric(data2$Tech_rep)

data2

# Refactoring COlumns for key
key$Sample_ID <- as.factor(key$Sample_ID)
key$Animal <- as.factor(key$Animal)
key$Condition <- as.factor(key$Condition)

key

# Refactoring columns for standards
standards2$Sample_ID <- as.factor(standards2$Sample_ID)
standards2$When <- as.factor(standards2$When)
standards2$Dilution_factor <- as.numeric(standards2$Dilution_factor)
standards2$Injection <- as.factor(standards2$Injection)
standards2$Nano_day <- as.numeric(standards2$Nano_day)

standards2

```



###Back calculate standards
```{r}
standards2 <- standards2 %>% 
  mutate(True_Count=Dilution_factor*Count)

# Set the correct order of 'categorical factors'
standards2$Nano_day <-  factor(standards2$Nano_day, levels=c('1'))
standards2$When <- factor(standards2$When, levels=c('before','after'))

standards2
```

###Summarize three technical standard replicates
```{r}
standards3 <- standards2 %>% 
  group_by(particle_size,Sample_ID,When,Dilution_factor,Nano_day,Injection) %>% 
  summarise( tech_N = length(True_Count),
             tech_mean = mean(True_Count),
             tech_sd = sd(True_Count),
             tech_se = tech_sd/sqrt(tech_N))
standards3
```


###Summarize standards by injection
```{r}
standards4 <- standards3 %>% 
  group_by(Nano_day,When,particle_size) %>% 
  summarise( inj_N = length(tech_mean),
             inj_mean = mean(tech_mean),
             inj_sd = sd(tech_mean),
             inj_se = inj_sd/sqrt(inj_N))
standards4

```

###Plot before and after plots, facet by experimental day
```{r}
std_plot <- standards4 %>% 
  ggplot(aes(x=particle_size,y=inj_mean,color=When))+
  geom_ribbon(aes(ymin=inj_mean-inj_se, ymax=inj_mean+inj_se),
              alpha=0.2,fill = alpha('grey12', 0.2)) + #error bars
  geom_line(size=2) + xlim(0,500)+ #line size, x-axis scale
  scale_y_continuous(expand=c(0,0))+ #set bottom of graph
  xlab("Particle Size") + # X axis label
  ylab("\nMean Particle Concentration/ml\n") + # Y axis label
  ggtitle("Nanosight Histogram of\nVirgin Mouse Plasma")+ #title
  labs(color="Condition")+ #Label table title
  facet_grid(. ~ When)

std_plot

```

### Standards particle concentrations from each experimental day
```{r}
standards_df <- standards4 %>% 
  group_by(Nano_day,When) %>% 
  summarise(total=sum(inj_mean))

standards_df
```

###Bar graph of standards particle concentrations
```{r}
standards_df %>% 
  ggplot(aes(x=Nano_day,y=total,fill=When))+
  geom_col(position="dodge")+
  scale_y_continuous(expand=c(0,0))+ #set bottom of graph
  xlab("Experimental Day") + # X axis label
  ylab("\nMean Particle Concentration/ml\n") + # Y axis label
  ggtitle("Nanosight Histogram of\nVirgin Mouse Plasma")+ #title
  labs(color="When") #Label table title
```

###Intraassay variability
```{r}
Intra.assay_cv <- standards_df %>% 
  group_by(Nano_day) %>% 
  summarise(Day_N = length(total),
             Day_mean = mean(total),
             Day_sd = sd(total),
             Day_se = Day_sd/sqrt(Day_N),
            Day_cv = Day_sd/Day_mean )
Intra.assay_cv
```

#Sample analysis

###Back calculate the original concentration of the sample
```{r}
data2 <- data2 %>% 
  mutate(True_Count = Dilution_factor*Count)
data2
```


### Average three technical readings
```{r}
data3 <- data2 %>% 
  group_by(particle_size,Sample_ID,Dilution_factor,Injection) %>% 
  summarise( tech_N = length(True_Count),
             tech_mean = mean(True_Count),
             tech_sd = sd(True_Count),
             tech_se = tech_sd/sqrt(tech_N))
data3
```

##Summarize samples by injection (average both injections)
```{r, warning=FALSE}

data4 <- data3 %>% 
  group_by(particle_size,Sample_ID,Dilution_factor) %>% 
  summarise( inj_N = length(tech_mean),
             inj_mean = mean(tech_mean),
             inj_sd = sd(tech_mean),
             inj_se = inj_sd/sqrt(inj_N))
data4

# Average technical replicates and merge with key
merge <- left_join(key,data3, by= "Sample_ID")

merge

# Average injection replicates and merge with key
merge1 <- left_join(key,data4, by= "Sample_ID")

merge1
```



#Quick visualizations

### Graphing all samples
```{r, warning=FALSE}
sample_plot <- merge %>%
  ggplot(aes(x=particle_size, y=tech_mean,color=Injection ))+ #plot
  geom_ribbon(aes(ymin=tech_mean-tech_se,
                  ymax=tech_mean+tech_se),
                  alpha=0.2,fill = alpha('grey12', 0.2)) + #error bars
  geom_line(size=2.0) + xlim(0,500)+ #line size, x-axis scale
  scale_y_continuous(expand=c(0,0))+ #set bottom of graph
  xlab("Particle Size") + # X axis label
  ylab("\nMean Particle Concentration/ml\n") + # Y axis label
  ggtitle("Nanosight Histogram of\nVirgin Mouse Plasma")+ #title
  labs(color="Injection")+ #Label table title
  facet_grid(Animal ~ Condition)+
  geom_vline(xintercept = 140)+
  annotate("text", x= 235, y = 1E9, label= "140nm")

sample_plot
```

### Particle concentration values for each of the samples
```{r}
merge2 <- merge1 %>% 
  group_by(Animal,Condition) %>% 
  summarise(particle_conc=sum(inj_mean))
merge2
```


### Summary statistics of particle concentration (averaging n=6 for each time point)
```{r}
merge3 <- merge2 %>% 
  group_by(Condition) %>% 
  summarise(Condition_N=length(particle_conc),
            Condition_mean = mean(particle_conc),
            Condition_sd = sd(particle_conc),
            Condition_se = Condition_sd/sqrt(Condition_N))
merge3
```

###Boxplot
```{r, warning=FALSE}
plot1 <- merge2 %>% 
  #filter(!Animal== '1371'|!Condition == 'lowOxygen') %>% 
  group_by(Condition) %>% 
  ggplot(aes(x= Condition, y = particle_conc, color=Condition)) +
  geom_boxplot(colour="black",fill=NA) + 
  geom_point(aes(text = paste("Animal:", Animal)),
             position='jitter',size=3)+
  xlab("\nTreatment\n") + # X axis label
  ylab("\nExosomes/ml\n") + # Y axis label
  ggtitle("GD 17.5 Placental Exosome \nExplant Culture (Ultracentrigution)\n")+ #title
  labs(color="Condition") # Label table title

  
plot1
```
##Interactive Plot
```{r}
ggplotly(plot1)
```


# Statistics

```{r}
fit <- t.test(particle_conc ~ Condition,data=merge2)

tidy(fit)
```

