---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

### Objective

Determine the optimal dose of Proteinase K treatment to remove outer membrane proteins off of GD14.5 placental EVs obtained by IZON size eclusion chromatography.

10ug of pEVs were treated with diffeing doses (10, 1, 0.1, 0.01 ug/ml) of Proteinase K in a volume of 20ul before incubation at 37C for 10 min.  Samples were then incubated at 70C for 5 minutes for ProK inactivation and 1ul of PMSF (protease inhibitor) was added.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyNano)
```

```{r}

df01_raw <- nanoimport("GD14_Pexo_ProK_titration_combined-ExperimentSummary.csv")

df01_raw
```

```{r}

samp_order <- c("PBS", "ProK001 ug/ml","ProK01 ug/ml","ProK1 ug/ml","ProK10 ug/ml")

df02_tidy <- df01_raw %>% 
  nanotidy(sep_var = c("GD", "Sample", "Treatment", "Dilution", "Tech_rep")) %>% 
  mutate(Treatment = str_replace_all(Treatment, "ugml", " ug/ml"),
         Treatment = factor(Treatment, levels = samp_order))

df02_tidy 
```
```{r message=FALSE, warning=FALSE}
Tech_rep_plot <- df02_tidy %>% 
  filter(particle_size < 701) %>% 
  ggplot(aes(particle_size, True_count, color = Tech_rep)) +
  geom_line(size = 1) +
  facet_wrap(~Treatment, nrow = 1) +
  themesoo::theme_soo() +
  labs(title = "Proteinase K treatment on IZON Pexos\n",
       y = "Concentration (Particles/ml)\n",
       x = "Particle size (nm)") 

Tech_rep_plot

# ggsave(plot = Tech_rep_plot, "Fig01_ProK_Titration_GD_Pexo_Tech_rep_plot.png", width = 10, height = 4, dpi = 600, units = "in")

```



```{r message=FALSE, warning=FALSE}
Dose_plot <- df02_tidy %>% 
  nanolyze(particle_size, GD, Sample, Treatment, param_var = True_count) %>% 
  ggplot(aes(particle_size, Param_mean, color = Treatment)) +
  geom_line(size = 1.5, alpha = 0.8) +
  themesoo::theme_soo() +
  labs(title = "Proteinase K treatment on IZON Pexos\n",
       y = "Concentration (Particles/ml)",
       x = "Particle size (nm)")

Dose_plot


# ggsave(plot = Dose_plot, "Fig02_ProK_Titration_GD_Pexo_mean_plot.png", width = 7, height = 5, dpi = 600, units = "in")
```

```{r message=FALSE, warning=FALSE}
count_plot <- df02_tidy %>% 
  nanocount(GD, Sample, Treatment,Tech_rep, param_var = True_count) %>% 
  ggplot(aes(forcats::fct_rev(Treatment), Particle_count , color = Treatment)) +
  geom_boxplot(outlier.colour = NA, alpha = 0.9) +
  geom_jitter(size = 4, width = 0.1, alpha = 0.7) +
  coord_flip() +
  labs(title = "Proteinase K treatment on IZON Pexos\n",
       caption = "Dots represent technical replicates",
       y = "Concentration (Particles/ml)\n",
       x = "Treatment") +
  guides(color = FALSE) +
  themesoo::theme_soo(title_size = 18)

count_plot



# ggsave(plot = count_plot, "Fig03_ProK_Titration_GD_Pexo_concentration_plot.png", width = 6, height = 4, dpi = 600, units = "in")
```

