---
title: "qEV Notebook"
author: "Sean Nguyen"
output:
  html_document:
    df_print: paged
---




```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(cowplot)
library(broom)
```

```{r message=FALSE, warning=FALSE}

data_path <- "data"   # path to the data
files <- dir(data_path, pattern = "*.csv$") # get file names


my_data <- data_frame(filename = files) %>% # create a data frame holding the file names
  mutate(file_contents = map(filename, 
                             ~ read_csv(file.path(data_path, .)))) # read files into a new data column  
```

```{r}
quick_summary <- function(df, ..., param_var) {
param_var <- enquo(param_var)
df %>%
group_by_(.dots = lazyeval::lazy_dots(...)) %>%
summarise(N = length(!!param_var),

mean = mean(!!param_var, na.rm = TRUE),
sd = sd(!!param_var, na.rm = TRUE),
se = sd/sqrt(N))

}
```


```{r}
raw_data <- my_data %>%
  filter(filename == "nanosight_data.csv") %>% 
  unnest() %>% 
  select(-filename)


data <- raw_data %>% 
  gather(samples,count,2:19) %>% 
  separate(samples, into = c("GD", "Aliquot","Dilution", "Injection", "Tech_rep")) %>% 
  mutate_at(vars(GD,Aliquot,Injection,Tech_rep),as.factor) %>% 
  mutate_at(vars(Dilution),as.numeric) %>% 
  mutate(true_count = Dilution * count)

data

```


```{r}
data %>%
  ggplot(aes(x = particle_size, y = true_count, color = Tech_rep)) +
  geom_line()+
  facet_wrap(GD~Aliquot)

```


# qEV of Plasma
  Plasma from a GD 16.5 mg x WT female was collected and analyzed by **Total Exosome Isolation (TEI) reagent [Thermo Fisher]** and by **qEV size exclusion columns [IZON]**.  100ul of plasma was used as input for each method.  
  
###TEI
  100ul of plasma was placed in a microfuge tube along with 50ul of PBS and 30ul of TEI reagent and subjected for centrifugation after a 10 minute incubation, the pellet was resuspended in 100ul of PBS.
  
###qEV 
  100ul of plasma was placed in a qEV column and 0.5ml fractions were collected, the void volume(3ml) was discarded and the subsequent 6th, 7th and 8th column of purified vesicles were selected.
  
### Nanosight analysis
  
  |Parameter|Value|  
  |---|---|  
  |Gain|1|
  |Level|12|  
  |Threshold|4|
  
```{r}
plasma_data <- my_data %>%
  filter(filename == "plasma_nanosight_data.csv") %>% 
  unnest() %>% 
  select(-filename)

plasma_data
```
  
```{r}
plasma_data_clean <- plasma_data %>% 
  gather(sample, value, 2:13) %>% 
  separate(sample, into = c("Sample", "ID", "Dilution", "filter","Tech_rep", sep = "_")) %>% 
  mutate_at(vars(Sample:ID,filter,Tech_rep),as.factor) %>% 
  mutate_at(vars(Dilution),as.numeric) %>%
  select(-(`_`)) %>% 
  mutate(True_count = Dilution * value)

plasma_data_clean
```
  
```{r}
plasma_data_clean %>% 
  # filter(!Sample == "TEI") %>% 
  ggplot(aes(x = particle_size, y = True_count, color = Tech_rep)) +
  geom_line() +
  facet_wrap(~Sample)
  
```
```{r}
plasma_data_clean %>% 
  quick_summary(Sample, ID, particle_size, param_var = True_count) %>% 
  ggplot(aes(x = particle_size, y = mean, color = Sample)) +
  geom_ribbon(aes(ymin = mean-se,
                  ymax= mean+se),
                  alpha=0.2,fill = alpha('grey12', 0.2)) + #error bars
  geom_line() +
  facet_wrap(~Sample, scales = "free_y")
```
  
```{r}
plasma_data_clean_summary <- plasma_data_clean %>% 
  quick_summary(Sample, ID, particle_size, param_var = True_count) %>% 
  group_by(Sample,ID) %>% 
  summarise(Total_conc = sum(mean)) %>% 
  mutate(Volume = case_when(
                    str_detect(Sample, "qEV") ~ "0.500",
                    str_detect(Sample, "TEI") ~ "0.100") %>% as.numeric(),
         Method = case_when(
                    str_detect(Sample,"qEV") ~ "qEV",
                    str_detect(Sample, "TEI") ~ "TEI") %>% as.factor(),
         Total_count = Total_conc * Volume)
plasma_data_clean_summary

```

```{r}
plasma_data_clean_summary %>% 
  group_by(Method) %>% 
  summarise(Total_count_by_method = sum(Total_count)) %>% 
  print() %>% 
  ggplot(aes(Method, Total_count_by_method, fill = Method))+
  geom_col() +
  scale_y_continuous(expand = c(0,0)) +
  labs( title = "Total number of particles by isolation technique\n",
        x = "Method",
        y = "Total number of particles")
```
  
  